---
title: "Initial Korean covid 19"
author: "Amber Walker"
date: "06/21/2023"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?

## Methods
We will analyze the sex of patients and the result of deceased of if they re alive or dead. I will use the variable in my collection to Cat-cat. This is a problem that we can use the helper file to assist in solving, after copying it into our project to get the correct answer.



## Graphical results - use ggplot commands to do this  

```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

  

There is more females tha m ales that show on these charts. Twice the amount of people that are males are dead as there is females, which is pretty odd since there are more females than men. it is killing males on a 2 to 1 basis. In or graphical results, Covid wants to kills more of the males than females.

  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
There is a total of 100 men and women each. The males that are dead is a percentage of 64, while the females haev a percentage of 36. The percentage of females that are dead is half of the percentage that are males.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
In the group of 1831,  were alive and 1411. All in all more females are alive than male. the P value is 1 out of a thousand. It is a small chance to get a female and a male dying. In Fishers test,9 pit of 10,000 are about the same p value. You can also see the odds ratio of the odds of female living against the odds of male dying. We want to know whats the odds of this is happening. 


## Conclusions

In today lesson, I learned how to do a ggplot with the helper file. I changed the class for result and the name of the project kRdata. I was also able to create a numerical graph in percentage and counting. I also ran a test and a fishers test and both are equal in p value. All in all the males are the ones who suffered more during this study.


